\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-15 create-FSHybridPLS.Rmd
code: get_xi_direct unified
remove get_xi_hat_linear

### helper: get_xi_hat_linear
```{r}
get_xi_hat_linear <- function(W, y){
  n <- W$n_sample
  K <- W$n_functional
  d = list()
  for (i in 1:K){  ## tPhi_tC : C multiplied by Phi. Scalable to arbitrary K
    Ct <- W$functional_list[[i]]$coefs
    Phi <- W$gram_list[[i]]
    rhs <-  t(Phi) %*% Ct %*% y
    d[[i]] <- solve(Phi, rhs)
  }
  d[[K+1]] <- t(W$Z) %*% y
  d_vec <- do.call(c, d)  # or: unlist(d)
  return(d_vec)
}
```

@@

>>>>

@@

# --- Step 1: Simulate Data ---
set.seed(111)

### 2025-07-15 manuscript/main.tex
remove get_xi_hat_linear

	We now describe the strategy for computing the PLS component direction at the $l$-th iteration. For simplicity, we omit the iteration index and assume that the observations have been residualized and are orthogonal to those from previous steps (see Section~\ref{subsec: Step 2: Computing Predictors via Hybrid-on-Scalar Regression} for the residualization procedure).
	The goal is to find a unit-norm direction $\xi \in 
	(\mathcal{B}^K \times \mathbb{R}^p) \subset \mathcal{H}$ that maximizes the squared empirical covariance between the resulting PLS component scores $\langle \widetilde{W}_1, \xi \rangle_{\mathcal{H}}, \ldots, \langle \widetilde{W}_n, \xi \rangle_{\mathcal{H}}$ and the responses $y_1, \ldots, y_n$,
	 defined as
\begin{equation}\label{def:squared_empirical_cov}
	 	\widehat{\textnormal{Cov}}^2 \left( \langle \widetilde{\mathbf{W}}, \xi \rangle_{\mathcal{H}}, Y \right)
	 	:=
	 	\biggl(
	 	\frac{1}{n} 
	 	\sum_{i=1}^n
	 	y_i
	 	\langle \widetilde{W}_i, \xi \rangle_{\mathcal{H}}
	 	\biggr)^2
	 	.
	 \end{equation}
	We denote the maximizer as
	\begin{equation}\label{def:maximizer_squared_empirical_cov}
		\hat{\xi} 
= 
(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\zeta}})
:= \arg \max_{\xi \in (\mathcal{B}^K \times \mathbb{R}^p) \subset \mathcal{H}} \widehat{\textnormal{Cov}}^2 \left( \langle \widetilde{\mathbf{W}}, \xi \rangle_{\mathcal{H}}, Y \right)
	\end{equation}
%
	where each functional part $\hat{\xi}_j $  is expressed as
	\begin{equation*}
		\hat{\xi}_j(t) := 
		\sum_{m=1}^M \hat{d}_{jm} \, b_m(t), \quad j = 1, \ldots, K.
	\end{equation*}

@@

	Define the Lagrangian:
	\[
	\mathcal{L}(\{\mathbf{d}_j\}, \boldsymbol{\zeta}, \lambda) = f(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) - \lambda g(\{\mathbf{d}_j\}, \boldsymbol{\zeta}).
	\]
	The constraint function $g$ is continuously differentiable, and its gradient vanishes only at the origin. Since the origin is not feasible (the constraint requires norm 1), the gradient of $g$ is nonzero at all feasible points. Thus, the linear independence constraint qualification (LICQ) holds, ensuring that the KKT conditions are necessary for local maxima (see, for example, Theorem 5.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}).
	Taking gradients of $\mathcal{L}$ and setting them to zero yields:
	\begin{align*}
		\nabla_{\mathbf{d}_j} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		B \Theta_j^\top \mathbf{y}
		- 2\lambda B \mathbf{d}_j = 0, \quad \forall j, 
		\numberthis \label{lag_grad_d}
		\\
		\nabla_{\boldsymbol{\zeta}} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		Z^\top \mathbf{y}
		- 2\lambda \boldsymbol{\zeta} = 0.
		\numberthis \label{lag_grad_zeta}
	\end{align*}
	By pre-multiplying equation \eqref{lag_grad_d} by $B^{-1}$ and observing that the scalar quantity $\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}$ appears in both \eqref{lag_grad_d} and \eqref{lag_grad_zeta}, the gradient conditions together imply that any local optimum must be  scalar multiples of
	$$
	\Theta_1^\top \mathbf{y}, \ldots, \Theta_K^\top \mathbf{y}, Z^\top \mathbf{y},
	$$
	all scaled by the same factor. There exists only one set of vectors  that also satisfies the primal equality constraint, namely the normalized version of the vector set above. The global maximizer is therefore given by
	\begin{equation*}
		\mathbf{d}_j^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		\Theta_j^\top \mathbf{y},
		\quad
		%%%%%%%%%%%%%%%%
		\boldsymbol{\zeta}^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		Z^\top \mathbf{y},
	\end{equation*}
	which are the same as  $\hat{\mathbf{d}}_1, \ldots, \hat{\mathbf{d}}_K$ and $\hat{\boldsymbol{\zeta}}$. This completes the proof of  Proposition \ref{proposition:generalized_rayleigh_quotient_linear_system_unregularized}.

>>>>

	We now describe the strategy for computing the PLS component direction at the $l$-th iteration. For simplicity, we omit the iteration index and assume that the observations have been residualized and are orthogonal to those from previous steps (see Section~\ref{subsec: Step 2: Computing Predictors via Hybrid-on-Scalar Regression} for the residualization procedure).
	The goal is to find a unit-norm direction $\xi \in 
	(\mathcal{B}^K \times \mathbb{R}^p) \subset \mathcal{H}$ that maximizes the squared empirical covariance between the resulting PLS component scores $\langle \widetilde{W}_1, \xi \rangle_{\mathcal{H}}, \ldots, \langle \widetilde{W}_n, \xi \rangle_{\mathcal{H}}$ and the responses $y_1, \ldots, y_n$,
	 defined as
\begin{equation}\label{def:squared_empirical_cov}
	 	\widehat{\textnormal{Cov}}^2 \left( \langle 
	 	\widetilde{W},
	 	\xi \rangle_{\mathcal{H}}, Y \right)
	 	:=
	 	\biggl(
	 	\frac{1}{n} 
	 	\sum_{i=1}^n
	 	y_i
	 	\langle \widetilde{W}_i, \xi \rangle_{\mathcal{H}}
	 	\biggr)^2
	 	.
	 \end{equation}
	We denote the maximizer as
	\begin{equation}\label{def:maximizer_squared_empirical_cov}
		\hat{\xi} 
= 
(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\zeta}})
:= \arg \max_{\xi \in (\mathcal{B}^K \times \mathbb{R}^p) \subset \mathcal{H}} \widehat{\textnormal{Cov}}^2 
\left( 
\langle \widetilde{W}, \xi \rangle_{\mathcal{H}}, Y 
\right)
	\end{equation}
%
	where each functional part $\hat{\xi}_j $  is expressed as
	\begin{equation*}
		\hat{\xi}_j(t) := 
		\sum_{m=1}^M \hat{d}_{jm} \, b_m(t), \quad j = 1, \ldots, K.
	\end{equation*}

@@

	Define the Lagrangian:
	\[
	\mathcal{L}(\{\mathbf{d}_j\}, \boldsymbol{\zeta}, \lambda) 
	:= 
	f(\{\mathbf{d}_j\}, \boldsymbol{\zeta}) - \lambda g(\{\mathbf{d}_j\}, \boldsymbol{\zeta}).
	\]
	The constraint function $g$ is continuously differentiable, and its gradient vanishes only at the origin. Since the origin is not feasible (the constraint requires norm 1), the gradient of $g$ is nonzero at all feasible points. Thus, the linear independence constraint qualification (LICQ) holds, ensuring that the KKT conditions are necessary for local maxima (see, for example, Theorem 5.3.1 in \citealt{bazaraaNonlinearProgrammingTheory2006}).
	Taking gradients of $\mathcal{L}$ and setting them to zero yields:
	\begin{align*}
		\nabla_{\mathbf{d}_j} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		B \Theta_j^\top \mathbf{y}
		- 2\lambda B \mathbf{d}_j = 0, \quad \forall j, 
		\numberthis \label{lag_grad_d}
		\\
		\nabla_{\boldsymbol{\zeta}} \mathcal{L} &= 
		\frac{2}{n^2} 
		\left(
		\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}
		\right)
		Z^\top \mathbf{y}
		- 2\lambda \boldsymbol{\zeta} = 0.
		\numberthis \label{lag_grad_zeta}
	\end{align*}
	By pre-multiplying equation \eqref{lag_grad_d} by $B^{-1}$ and observing that the scalar quantity $n^{-2}\sum_{\ell=1}^K \mathbf{y}^\top \Theta_\ell B \mathbf{d}_\ell + \mathbf{y}^\top Z \boldsymbol{\zeta}$ appears in both \eqref{lag_grad_d} and \eqref{lag_grad_zeta}, the gradient conditions together imply that any local optimum must be  scalar multiples of
	$
	\Theta_1^\top \mathbf{y}, \ldots, \Theta_K^\top \mathbf{y}, Z^\top \mathbf{y},
	$
	all scaled by the same factor. There exists only one set of vectors  that also satisfies the primal equality constraint, namely the normalized version of the vector set above. The global maximizer is therefore given by
	\begin{equation*}
		\mathbf{d}_j^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		\Theta_j^\top \mathbf{y},
		\quad
		%%%%%%%%%%%%%%%%
		\boldsymbol{\zeta}^\ast = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		Z^\top \mathbf{y},
	\end{equation*}
	which are the same as  $\hat{\mathbf{d}}_1, \ldots, \hat{\mathbf{d}}_K$ and $\hat{\boldsymbol{\zeta}}$. This completes the proof of  Proposition \ref{proposition:generalized_rayleigh_quotient_linear_system_unregularized}.
